---
title: "Lab 9"
author: "Sophia Leiker"
date: "11/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objects and classes

## Vectors

### Creating a vector

```{r}
#numeric vector
my_vector <- vector(mode = "numeric", length = 100)
#logical vector
my_second_vector <- vector(mode = "logical", length = 20)
```

### Vector propoerties

#### Length

```{r}
# length of my vector
length(my_vector)
length(my_second_vector)
```

#### Class

```{r}
#what's in my vactor?
class(my_vector)
class(my_second_vector)
```

#### Indexing (assessing vectors)

```{r}
# Using c to create a vector
my_vector <- c(1:50)

#What's the first element?
my_vector[1]

#What's the last element?
my_vector[-1]

# What's my seventh element 
my_vector[7]

#Select a range from a vector?
my_vector[3:7]
```

## Matricies

### Creating matricies

```{r}
# Create a 10 X 10 matrix full of NA's
my_matrix <- matrix(data = 1:100, nrow = 20, ncol = 5)
```

### Matrix propoerties

#### Dimensions

```{r}
#What are the dimensions

#The first output is the number of rows, the second value is the number of columns
dim(my_matrix)
```

#### Length

```{r}
#This means there are 100 elements (if to stretch it into a line, you see that there are 100 values)
length(my_matrix)
```

#### Class
```{r}
class(my_matrix)
```

#### Indexing
First number indicates rows
```{r}
#This goes to row two and gives all the elements (gives all the values in the second row)
my_matrix[2,]

# Index a column (extracting third column from the matrix)
my_matrix[,3]

# To index a given column in 8th row, column 3
my_matrix[8,3]

# Index and replace a value in row 8 column 3
my_matrix[8,3] <- 100000

# Replace an entire column
my_matrix[,5] <- 10
```

# Control flow

## `for` loops

variable (counter), sequence, expression

```{r}
for (i in 1:10) {
  print(i)
}

for (i in 1:10) {
  print("Hello world!")
}
```

# Generating random numbers
Using rnorm
```{r}
#Every time you run this, you will get 5 different values for the specified requirements outlined below
rnorm(n =5, mean = 10, sd = 2)
```

# Let's combine all of the above 

We will use a `numeric` to create an empty vector
We will use `nfor` to interate along items of that emptry vector
We will use [] to access each slot and fill-in that empty vector
We will use rnorm to sample values from a random normal distribution
We will plot the trend 

$$N_(t) =N_{t-1} + r\left(1- \frac(N_{t-1}{K}\right))$$

## Writing the code down

```{r}
#Set up model parameters
K <- 100
nsteps <- 100
nu <- 0.4
sigma <- 1

# Set up the model objects
N <- vector(mode = "numeric", length = nsteps)
N[1] <- 25

#Model it
for(t in 2:nsteps){
  r <- rnorm(n=1, mean = nu, sd = sigma)
  N[t] <- N[t-1] + r * (1 - (N[t-1] /K))
}

# Visualize it
plot(N)

```

# Somthing more
We will use matrix to create an emptry matrix (rows are time, columns are runs)
We will repeat the process

```{r}
#Set up model parameters
K <- 100
nsteps <- 100
nu <- 0.4
sigma <- 1

# Set up the model objects
N <- matrix(data = NA, nrow = nsteps, ncol = 100)
N[1,] <- 25

#Model it
for(t in 2:nsteps){
  r <- rnorm(n=100, mean = nu, sd = sigma)
  N[t,] <- N[t-1,] + r * (1 - (N[t-1,] /K))
}

# Visualize it
#atplot(N, type = 1) isn't working
```

# Bring in your friends!

Convert it to ggplot

```{r}
#load tidyverse
library(tidyverse)

#create data.frame
my_data <- as_tibble(N) %>% 
  mutate(time = 1:nsteps) %>% 
  gather(run, N, -time)

# Visualize it
ggplot(data = my_data, mapping = aes(x = time, y = N)) +
  geom_line(aes(group = run)) +
  stat_summary(geom = "line", fun = mean, color = 'red') +
  theme_bw() +
  labs(x = "Time (years)")
```


